### 2015-08-13 abstract.tex
new figure

Scientific discovery at the exascale will not be possible without significant
new research into the management, storage and retrieval over the full lifespan
of the extreme amounts of data that will be produced. Our thesis is that by
adding application level knowledge about data to guide the actions of the
storage system we obtain substantial benefits in the organization, storage, and
access to extreme scale data, resulting in improved productivity for
computational science. We will demonstrate novel techniques to facilitate
efficient mapping of data objects from the user space onto multiple storage
tiers, and enable application-guided data reductions and transformations to
address capacity and bandwidth bottlenecks.  Our goal is to address the
associated Input/Output (I/O) and storage challenges in the context of current
and emerging storage landscapes, and expedite insights into mission critical
scientific processes.  We aim to reduce the time to insight, not just for a
single application, but for the entire workload in a multi-user environment,
where storage is shared across users. We shall achieve this by allowing
selectable data quality, through trading in accuracy and error in order to meet
the time or resource constraint. We will explore beyond checkpoint/restart I/O,
and will address the challenges posed by key data access patterns in the
knowledge gathering process. Ultimately, we will take the knowledge from the
storage system to provide vital feedback to the middleware so that the best
possible decisions can be autonomically made between the user intentions and
the available system resources.  We will test our prototypes on current and
future Department of Energy (DOE) system with many of today's cutting edge
applications to ensure our techniques can be used within our framework on
current and next generation systems.

>>>>

% Scientific discovery at the exascale will not be possible without significant
% new research into the management, storage and retrieval over the full lifespan
% of the extreme amounts of data that will be produced. Our thesis is that by
% adding application level knowledge about data to guide the actions of the
% storage system we obtain substantial benefits in the organization, storage, and
% access to extreme scale data, resulting in improved productivity for
% computational science. We will demonstrate novel techniques to facilitate
% efficient mapping of data objects from the user space onto multiple storage
% tiers, and enable application-guided data reductions and transformations to
% address capacity and bandwidth bottlenecks.  Our goal is to address the
% associated Input/Output (I/O) and storage challenges in the context of current
% and emerging storage landscapes, and expedite insights into mission critical
% scientific processes.  We aim to reduce the time to insight, not just for a
% single application, but for the entire workload in a multi-user environment,
% where storage is shared across users. We shall achieve this by allowing
% selectable data quality, through trading in accuracy and error in order to meet
% the time or resource constraint. We will explore beyond checkpoint/restart I/O,
% and will address the challenges posed by key data access patterns in the
% knowledge gathering process. Ultimately, we will take the knowledge from the
% storage system to provide vital feedback to the middleware so that the best
% possible decisions can be autonomically made between the user intentions and
% the available system resources.  We will test our prototypes on current and
% future Department of Energy (DOE) system with many of today's cutting edge
% applications to ensure our techniques can be used within our framework on
% current and next generation systems.

This project explores the use of application level knowledge to reduce the
time to insight for a workload (not just a single application) in a
multi-user environment where storage and other resources are shared among
users. Our approach is based on the notion of selectable data quality to
enable a tradeoff between result accuracy and near-real-time data movement
requirements for oversubscribed computational and storage resources. For the
past several decades, extreme scale computational science has been conducted
on dedicated computational and storage resources. Shared resources were
overprovisioned resulting in little or no impact from resource contention.
This changed recently. Execution times for scientific applications vary
dramatically depending on external contention for shared resources. This
execution time variance is particularly challenging in light of new usage
models that have near-real-time requirements, e.g., providing information
needed to control an experimental instrument like the Spallation Neutron
Source.

More specifically, we will explore the resource contention issues
with spreading a single data set horizontally across a single tier in the
storage hierarchy, such as spinning media, as well as vertically across
different kinds of storage media, including non-volatile memory. To manage
the varying storage media capacities and bandwidths, we provide mechanisms
to express the required data quality and the near real-time requirements of
the application. Our thesis is that by adding application level knowledge
about data to guide storage system behaviors, we obtain substantial benefits
in the organization, storage, and access to extreme scale data resulting in
improved productivity for computational science. We will demonstrate novel
techniques to facilitate efficient and effective data placement onto
multiple storage tiers and enable application-guided data reductions and
transformations to address capacity and bandwidth bottlenecks. Our goal is
to address the associated data management challenges in the context of
current and emerging storage landscapes and expedite insights into mission
critical scientific processes
